Approximation as Differentiation

### 2025-05-08 main/formal-structures.tex
Grothendieck.
Galois slicing via Grothendieck.

\begin{definition}[Reindexing]
For any function $f: X \to Y$ define the \emph{reindexing} functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to
\Func{X}{\cat{C}}$ which sends any $Y$-indexed family $F$ over $\cat{C}$ to the $X$-indexed family
$\reindex{F}{f} = F \comp f$, with $f$ interpreted as the unique functor between $X$ and $Y$ as discrete
categories.
\end{definition}

@@

>>>>

\begin{definition}[Reindexing]
For any function $f: X \to Y$ define the \emph{reindexing} functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to
\Func{X}{\cat{C}}$ which sends any $Y$-indexed family $F$ over $\cat{C}$ to the $X$-indexed family
$\reindex{F}{f} = F \comp f$, regarding $f$ as a functor between $X$ and $Y$ as discrete categories.
\end{definition}

@@

\begin{definition}
\label{def:grothendieck:Grothendieck}
Suppose an indexed category $T: \cat{C}^\op \to \Cat$. The \emph{Grothendieck construction} for $T$ is the
category $\Grothendieck{X}T(X)$ (also written $\Grothendieck{}T$) which has as objects, all pairs $(X,
\partial X)$ of an object $X$ of $\cat{C}$ and an object $\partial X$ of $T(X)$, and as morphisms $(X,
\partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of morphisms $f: X \to Y$ in $\cat{C}$ and
morphisms $\partial f: \partial X \to T(f)(\partial Y)$ in $T(X)$.
\end{definition}

\noindent The $\partial X$ is intended to suggest the connection between the Grothendieck construction and the
idea of tangent spaces and derivatives. To understand how the $\partial X$ components compose, consider any
$(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ and $(g, \partial g): (Y, \partial Y) \to (Z, \partial
Z)$ in $\Grothendieck{}T$. We have:

\begin{itemize}
\item $\partial f: \partial X \to T(f)(\partial Y)$;
\item $T(f)(\partial g): T(f)(\partial Y) \to T(f)(T(g)(\partial Z)) = T(g \comp f)(\partial Z)$.
\end{itemize}

\noindent where $T(f)(\partial g)$ is the image of $\partial g$ in the functor $T(f)$. The composition $(g,
\partial g) \comp (f, \partial f): (X, \partial X) \to (Z, \partial Z)$ is then given by $(g \comp f,
T(f)(\partial g) \comp \partial f)$.

\subsubsection{Galois slicing via the Grothendieck construction}

Now consider the total category $\Fam(\LatGal)$, which is our basic setting for interpreting first-order
programs. This has as objects $(X, \partial X)$, all pairs of a set $X$ and and for every $x \in X$, a bounded
lattice $\partial X_x$, and as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of
a function $f: X \to Y$ and for every $x \in X$, a Galois connection $\partial f(x): \partial X(x) \to
\partial Y(f(x))$, as required. The role of the pullback (reindexing) along $f$ in the composition $(g,
\partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f)$ is to select the
appropriate lattice of approximations and Galois connection at each point, using the baseline (unapproximated)
function $f$.
